---
title: "QRM II Graded Assignment (7)"
subtitle: "Period 1, 2026"
author: "K.Stokvis 2859043, M.Braet 2860223, O.Keijser 2862760, Z.Cheung 2802886, K.Hendriks 289447"
date:  "`r format(Sys.Date(), '%d-%m-%Y')`"
output:
  pdf_document:
    number_sections: true
  html_document:
    df_print: paged

header-includes:
- "\\setcounter{section}{-1}"
- |
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
    showspaces = false,
    showtabs = false,
    breaklines,
    commandchars=\\\{\}
  }
format:
  html:
    code-overflow: wrap
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
movies3<-read.table("movies3_updated.tsv",sep="",header=TRUE)
library(dplyr)
library(ggplot2)
library(moments)
```

# Introduction

This assignment is to be completed in groups of 4-5. Further, all students in your group need to be assigned to the same R tutorial group (Friday's tutorial). You can sign yourself up for a group on Canvas. Please do so \textbf{before the start of your first R tutorial on Friday September 4th.} You can use the Discussion Board in Canvas if you do not have a group yet or if your group is incomplete.

The assignment has 5 parts, and each part corresponds to the course material of that week (with the exclusion of week 6, for which there is no R programming material). 

You are supposed to hand in these assignments on Canvas at the following dates:

  - **Deadline 1** *Sunday September 27th, at 23:59pm*: you are supposed to hand in weeks 1, 2, and 3 of this assignment.   This will determine 18\% of your overall course grade
  - **Deadline 2** *Sunday October 11th, at 23:59pm*: you are supposed to hand in weeks 4, and 5 of this assignment. This will determine 12\% of your overall course grade 

The R tutorials (each Friday) will consist of two halves. During the first half, you will discuss the tutorial exercises. These can be downloaded separately from Canvas. During the second half, you can work on this graded assignment within your own group. The purpose is that you find out how to work with R for doing statistical analyses by yourself. The tutorial exercises are meant to teach you basic commands to get you started, but to answer the problem sets in this assignment, you might need to research your own solutions, and use functions and commands not described in the tutorial exercises. Learning how to solve your own research problems is integral part of learning R. When you and your group get stuck on how to approach an exercise, the hierarchy in finding your way is as follows:

-	use the concepts from the tutorial exercises;
- use the cheat sheets available on Canvas;
-	use Google, YouTube, StackOverflow, or another website;
- ask the teacher.

The use of generative AI is \textbf{not} permitted and may result in a grade of 0. See the AI protocol in the course manual for details.

To answer the assignment, you can simply fill out this R markdown document. There are designated places which you can fill with R code. There are also designated spaces for you to answer each question. Often, the structure of an answer will be as follows. First, you type the R code in the designated box. This will show how you analyzed the data to get the answer to the question. Below the box for the R code, you will then summarize your answer to the question, i.e. what are the conclusions that you draw from the data analysis? 

When handing in, you are supposed to submit this .Rmd file, and a knitted version of this document. You can knit this document to pdf, word, or html. Knitting to pdf requires you to have a .tex distribution installed on your computer. Knitting to Word requires you to have Word installed. 

The exercises are designed such that you should be able to finish the majority of them during the tutorial each week. If you are not able to finish them fully during that time, you are expected to work on it in your own time using the computers on campus or your own device. It is best to meet as a group in-person when working together. If you want to work remotely, github is a good platform to guarantee smooth collaboration. Alternatively, you can email this .Rmd file back and forth to one another as a group, but this is not recommended as it is more cumbersome.

We encourage you to keep your code blocks, printing statements, and final answers, as short as possible. In any case, there is a page limit of 6 pages per week, which encompasses the total length of this document which consists of the questions, your coding lines, and your answers. When your answers to questions of the respective week exceed this page limit, they will not be graded, resulting in zero points.

Each week consists of 1, 2, or 3 subquestions. The total amount of points you can earn per week is 20 points. 

# Week 4 

1. There is an argument going on in the movie studio. *Bob* claims that production budgets are getting out of hand, and that the studio should focus on making cheaper movies. *Chantal* disagrees. She tells Bob that ``The profitability ratio of the movie will remain similar, no matter how much the budget is''. 

a. Set up a regression model to test Chantal's claim, and estimate it. That is, estimate: $$\text{PR}_i=\beta_0+\beta_1 \text{Budget}_i +\varepsilon_i.$$ Print a summary of your estimated model. \textbf{[2 points]}
b. What is the estimated value of $\beta_1$ how do you interpet it? \textbf{[2 points]}
c. Test for the null hypothesis that $\beta_1 = 0$. Report the p-value and state your conclusion. \textbf{[2 points]}
d. Who do you think is correct? Bob or Chantal? What would you advise the movie studio to do? \textbf{[2 points]}
e. Make a histogram of the residuals of your model. Based on the histogram, do you have any reason to doubt the validity of your model? Explain. \textbf{[2 points]}
f. Estimate hatvalues for each observation that you used to estimate your model. Do any hatvalues exceed the critical value? Based on this analysis, do you hvae any reason to doubt the validity of your model? \textbf{[2 points]}
g. Based on your analysis in f.), estimate your model again, but remove any outliers. Print a summary of your model and again test for the null hypothesis that $\beta_1 = 0$. What do you conclude? Based on this new model, would your advice to the movie studio in question d. change? \textbf{[2 points]}

----------------------------------------------------IMPORTANT---------------------------------------------------------
 **we uploaded the adjusted dataset "movies3" from the previous 3 weeks and called it "movies3_updated.tsv". The reason for this is that certain exercises for week 4 and week 5 needed variables from the previous week. An example of this would be the variable "PR" in week 4. We uploaded our movies3 dataset to this rmd file with the help of our TA**
----------------------------------------------------------------------------------------------------------------------

*step a*

```{r}
#WRITE YOUR CODE HERE
lm <- lm(PR~budget, data=movies3) 
summary(lm) 
``` 

**Your Answer:**

In the summary we can see the estimates and standard errors for beta0 and beta1. We can also check the fit of our estimated model with R-squared.

*step b*

```{r}
#WRITE YOUR CODE HERE
coef_budget <- summary(lm)$coefficients["budget", "Estimate"] 
print(coef_budget) 
```

**Your Answer:**

For each additional unit (euros or dollars) of budget, the Profitability Ratio (PR) decreases by -3.688924e-05. 

*step c*

```{r}
#WRITE YOUR CODE HERE
p_budget <- summary(lm)$coefficients["budget", 4] 
print(p_budget) 
```

**Your Answer:**

The p-value is 0.3737, even with very high significance levels $\alpha$, the p-value will still be larger than $\alpha$. Based on the fact that p-value > $\alpha$, we cannot reject the null hypothesis of $\beta_1 = 0$.

*step d*

```{r}
#WRITE YOUR CODE HERE
print(p_budget)
```

**Your Answer:**
Because we cannot reject that the slope is not zero, it means that a change in budget has no significant correlation with the Profitability Ratio. We therefore conclude that Chantal is correct. We would advise the movie studio not to decrease the budget of movies in order to increase the profitability ratio of the movies. Instead we advise the movie studio to look at other variables that might impact the profitability ratio. 

*step e*

```{r}
#WRITE YOUR CODE HERE
skewness <- skewness(lm$residuals)
print(skewness)
hist(residuals(lm), breaks = 5, 
    main = "Histogram of Residual Levels", 
    xlab = "Residuals")
```

**Your Answer:**

The residuals should be normally distributed, but when looking at the graph and at the skewness value of 23.44 we see that the residuals are strongly right-skewed. This raises concerns about the validity of our model.

*step f*

```{r}
#WRITE YOUR CODE HERE
n <- length(lm$model$PR) 
levcrit <- 2/n 
hatvalue <- hatvalues(lm) 
suspect <- length(hatvalue[hatvalue>levcrit]) 
print(suspect) 
sus_fraction = suspect/n 
print(sus_fraction) 
```

**Your Answer:**

66 out of the 555 observations are suspect and exceed the critical value, which is 11.89% of total observations.11.89% of all observations being an outlier is quite high and could negatively influence the estimators of b0 and b1. This gives reason to doubt the validity of the model. However, these observations are still real and should not just be omitted to create a better model. 

*step g*

```{r}
#WRITE YOUR CODE HERE
no_outliers <- hatvalue <= levcrit
lm_new <- update(lm, data = lm$model[no_outliers,])
p_newbudget <- summary(lm_new)$coefficients["budget", 4] 
print(p_newbudget)
summary(lm_new)

```

**Your Answer:**

We see that the estimators of beta0 and beta1 changed after we removed all 66 outliers that exceeded the critical value of n/2. We see that our p-value decreased from 0.37 to 0.23. This p-value is still too large to be smaller than any reasonable significance levels (alpha). Based on this observation we still cannot reject the null hypothesis. 

2.

a. Make a scatterplot with vote_average on the y-axis, and runtime on the x-axis. Within the same scatterplot, draw a straight line which summarizes the association between the average vote and the runtime. \textbf{[3 points]}
b. Using the insights from the scatterplot, your linear model, and your data, find an example of a movie that vastly overperformed. That means, find a movie that had much higher average vote than expected, given its runtime. For this movie, report its vote_average, its predicted vote_average, and its residual. \textbf{[3 points]}

*step a*

```{r}
#WRITE YOUR CODE HERE
library(ggplot2)

ggplot(
  subset(movies3,runtime !=0 & vote_average!=0), 
  aes(x = runtime, y = vote_average)) + 
  geom_point(na.rm=TRUE) + 
  geom_smooth(method = "lm", na.rm=TRUE) +
  scale_x_continuous(limits = c(55,205)) +
  scale_y_continuous(limits = c(0,10))
```

**Your Answer:**

We excluded any movies that had zero runtime and no vote average as these movies had no relevant data. Based on the scatterplot we see a potential positive correlation between vote average and runtime.

*step b*

```{r}
#WRITE YOUR CODE HERE

scatterplot_data<-subset(movies3,runtime !=0 & vote_average!=0)
reg_model<-lm(vote_average ~ runtime,movies3,subset = runtime !=0 & vote_average!=0)
scatterplot_data$vote_prediction<- predict(reg_model)
scatterplot_data$residual<-residuals(reg_model)
head(
  scatterplot_data[order((residuals(reg_model)), decreasing = TRUE), c("title", "runtime", "vote_average","vote_prediction","residual")], 1)

```

**Your Answer:**

We first created a new dataset, which excluded irrelevant observations (see a). Next, we made a regression model for determining the vote average based on the runtime. Using this model we created a new variable which predicted the average voting based on runtime. After this we added a variable for the residuals that showed the difference between actual vote average and predicted vote average. Finally we ordered the residual data with largest residual values first. Based on this data we can conclude that "The Lion King" overperformed its vote_average the most with 2.22 points.

# Week 5
  
a. Create a plot of the mean PR by year of release. Include 95\% confidence intervals around each mean Do you see any indication of a time trend in the PR? \textbf{[4 points]}
b. Estimate an OLS model which has as dependent variable the PR of a movie, and as independent variable the log of budget, a dummy for each level that genre2 can take, and the year of release. Show a summary of the resulting model and interpret each coefficient. \textbf{[4 points]}

The movie studio that you work is releasing a new movie in May of 2026. It will be a Comedy movie with a budget of 40,000,0000. It will have a runtime of 160 minutes. 

c. Estimate a model that is able to predict the revenue of this movie. Give its predicted revenue and include a 95% prediction interval. \textbf{[6 points]}
d. Plot the residuals against various independent variables in the model. Do you find any evidence of heteroskedasticity? \textbf{[2 points]}
e. The movie is set in Germany. The director wants to make the movie with only german-speaking actors. The executive producer is strongly against this, as she is afraid that movies that are not english spoken will have a low profit ratio. Estimate a model that can test whether the language spoken impacts the PR of this specific movie. Use an appropriate hypothesis test to test whether the language. What would you advise the movie studio regarding the language to be spoken? \textbf{[6 points]}


*step a*

```{r}
library(ggplot2)
movies3_clean <- movies3 %>%
  filter(PR < 4000, na.rm = TRUE)

sum(movies3$PR > 4000, na.rm = TRUE)

PR_year <- movies3_clean %>%
  group_by(release_year) %>%
  summarise(mean_PR = mean(PR, na.rm = TRUE),
            n = sum(!is.na(PR)),
            sd_PR = sd(PR, na.rm = TRUE))

PR_year_LU <- PR_year %>%
  mutate(se = sd_PR / sqrt(n),
            lower = mean_PR - qt(0.975, df = n - 1) * se,
            upper = mean_PR + qt(0.975, df = n - 1) * se)


ggplot(PR_year_LU,aes(x = release_year, y = mean_PR)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  scale_x_continuous(
    breaks = seq(
      min(PR_year_LU$release_year),
      max(PR_year_LU$release_year),
      by = 2))+
  coord_cartesian(ylim = c(-45,45))+
  labs(
    x = "Year of release",
    y = "Mean PR",
    title = "Mean PR by year of release") +
  theme_minimal()

```

**Your Answer:**

(In the plot shown above, two "outliers" are deleted from the data set, to be able to read the plot without difficulty. These are values that have a PR value above 4000.)

The plot does not clearly indicate any trend of the PR values over time.



*step b*

```{r}
length(unique(movies3$genre2))
unique(movies3$genre2)

movies3 <- movies3 %>%
  mutate(log_budget = log(budget))

movies3$genre2 <- as.factor(movies3$genre2)
movies3$genre2 <- relevel(movies3$genre2, ref = "Other")

model <- lm(PR ~ log_budget + genre2 + release_year, data = movies3)
summary(model)

```

**Your Answer:**
**outliers hier niet weggehaald, dat is nog een mogelijkheid, moeten de units er ook bij?**
(All interpretations are made under the ceteris paribus assumption.)
A increase in the budget with 1% will lower the PR value by -61.138. 
If the film has the genre of "Comedy" the PR value will be 2283.0 lower than if the film was of the categorie "Other" (default in this model). If the film has the genre "Drama" the PR will be 3121.0 lower than if the film was of the categorie "Other" and if the movie was of the genre "Thriller" than the PR value will be 9424.9 higher than if the movie was of the categorie "Other". 
If the movie is brought out one year later, the PR value will be 183.4 higher. 
We can also try to interpet the intercept, however then we would get a value for a movie that has a budget of one and is of the genre "Other" and is released in year zero, wich is not meaningfull because the year zero is outside of the relevant range of (our) data. 


*step c*

```{r}
# Step c: filter out missing budget/revenue (0 = unknown) and estimate the model
movies3_c <- movies3 %>%
  filter(budget > 0, revenue > 0)

model_c <- lm(revenue ~ log(budget) + genre2 + release_year + runtime, data = movies3_c)
summary(model_c)

# New movie
newbudget  <- c(40000000)
newgenre   <- c("Comedy")
newyear    <- c(2026)
newruntime <- c(160)

newdata <- data.frame(newbudget, newgenre, newyear, newruntime)
colnames(newdata) <- c("budget", "genre2", "release_year", "runtime")

predict(model_c, newdata, interval = "prediction")
```

**Your Answer:**

The estimated revenue is 289,051,538 dollar. The 95% prediction interval is very big the lower boundary is -9,381,514 and the upper boundary is 587,484,591 dollar. The lower boundary doesn't really make sense in reality, because revenue cannot be negative. For this question we used 40,000,000 as the budget and not 40,000,0000 because we assumed this was a typo.

*step d*

```{r}
plot(model_c$model$runtime, model_c$residuals,
xlab = "Runtime",
ylab = "Residuals",
main = "Residual Plot for model_c")

plot(model_c$model$release_year, model_c$residuals,
xlab = "Release Year",
ylab = "Residuals",
main = "Residual Plot for model_c")

plot(model_c$residuals ~ model_c$model$`log(budget)`, 
xlab = "log(budget)", 
ylab = "Residuals", 
main = "Residual Plot for model_c")

boxplot(model_c$residuals~model_c$model$genre2,
xlab = "Genre2",
ylab = "Residuals",
main = "Residual Plot for model_c")


```

**Your Answer:**

Runtime and release year seem to be pretty much homoskedastic. The variance of the log(budget) seems to increase when the log(budget) increases, so there is heteroskedasticity. The variance of other in genre2 is much larger than for the other genre2 categories. This also implicates heteroskedasticity.

*step e*

```{r}
#WRITE YOUR CODE HERE
#The movie is set in Germany. The director wants to make the movie with only german-speaking actors. The executive producer is strongly against this, as she is afraid that movies that are not english spoken will have a low profit ratio. Estimate a model that can test whether the language spoken impacts the PR of this specific movie. Use an appropriate hypothesis test to test whether the language. What would you advise the movie studio regarding the language to be spoken?

movies3$language_EN <- ifelse(movies3$original_language == "en",1,0)

model_e <- lm(formula = PR ~ budget + genre + release_year + runtime + language_EN, data = movies3)

newbudget  <- c(40000000)
newgenre   <- c("Comedy")
newyear    <- c(2026)
newruntime <- c(160)
english <- c(1,0)

newdata <- data.frame(newbudget, newgenre, newyear, newruntime, english)
colnames(newdata) <- c("budget", "genre", "release_year", "runtime", "language_EN")

coefficients <- summary(model_e)$coefficients

t_stat <- coefficients["language_EN", "t value"]
degrees_of_freedom <- df.residual(model_e)

p_value <- pt(t_stat, df = degrees_of_freedom, lower.tail = FALSE)

p_value


```
H_0: language_EN coefficient <= 0, H_1: language_EN coefficient > 0
We have chosen a righ tailed test because we want to test if an english spoken movie is expected to have a higher profitability ratio than a non-english spoken movie. 
significance level: 0.05 
The p value is calculated at 0.325 > 0.05 which means we reject the null hypothesis. 
It then follows that the spoken language of a movie has a scientifically significant effect on the profitability ratio. We advise the movie studio to use english as the spoken language.